---
title: "Stone Age Plague"
output: html_notebook
---

##Description Notebook
This notebook will contain any bits of analysis conducted in samples positive for plague during the Stone Age (and possibly later, only prehistorical genomes). 

#Checking branch OOH003-RISE505-ARS007-KZL002
```{r}
library(tidyverse)
snpTable_OOH003c <- read.delim("/projects1/pestis/lnba_paper_2020/vcfAnalysis/MS_3X_snps_falseSNPexcluded_LNBA_07072020/snpTable_OOH003_c.tsv")
noncalls <- c(".","N")

  snpTable_OOH003c %>%
  select(Position, OOH003_c, RISE505, ARS007, KZL002, KLE048) %>%
  filter(!OOH003_c %in% noncalls & RISE505 %in% noncalls & KLE048 %in% noncalls) %>%
  filter(OOH003_c == ARS007 | OOH003_c == KZL002)

snpTable_OOH003c %>%
  select(Position, OOH003_c, RISE505, ARS007, KZL002, KLE048) %>%
  filter(!RISE505 %in% noncalls & OOH003_c %in% noncalls & KLE048 %in% noncalls) %>%
  filter(RISE505 == ARS007 | OOH003_c == KZL002) 

snpTable_OOH003c %>%
  select(Position, OOH003_c, RISE505, ARS007, KZL002, KLE048) %>%
  filter(!RISE505 %in% noncalls & !OOH003_c %in% noncalls & !ARS007 %in% noncalls & !KZL002 %in% noncalls & KLE048 %in% noncalls)

snpTable_OOH003c %>%
  select(Position, OOH003_c, RISE505, ARS007, KZL002, KLE048) %>%
  filter(!RISE505 %in% noncalls & !OOH003_c %in% noncalls & ARS007 %in% noncalls & KZL002 %in% noncalls & KLE048 %in% noncalls)

##Branch RK1 and RISE509
snpTable_OOH003c %>%
  select(Position, RISE509, RK1001.C, GEN72, GRS004, Gyvakarai1, XXX001.A18962.69.72, GZL002.A0101_02.YP2.1, GZL001.A0101_02.YP2.1, KunilaII, HUT004.A_1343UnTal85, HOP001, HOP004, PST006.A_Post6, KLE031, KLE048, RISE505, OOH003_c, ARS007, KZL002) %>%
  filter(!RISE509 %in% noncalls & RK1001.C %in% noncalls & !GEN72 %in% noncalls)

snpTable_OOH003c %>%
  filter(HUT004.A_1343UnTal85 == KLZ001 & HUT004.A_1343UnTal85 != HOP001)
```

#Genetic vs geographical distance
###Package and function loading
```{r}
#Calculate pairwise genetic distance
library(tidyverse)
library(geosphere)
library(reshape2)
library(ggplot2)
library(ape)
library(plotly)
library(ggpubr)
###FUNCTIONS###

matrixToLong <- function(matrix, name, levels){
  longdf <- melt(matrix, value.name = name)
  longdf$Var1 <- factor(longdf$Var1, levels = levels)
  longdf$Var2 <- factor(longdf$Var2, levels = levels)
  return(longdf)
}

GenDistanceCalculator <- function(snpData, model, pairwise, levels, snpTable){
  percentDistance <- dist.dna(snpData, model = model, 
                              as.matrix = TRUE, 
                              pairwise.deletion = pairwise)
#https://rdrr.io/cran/ape/man/dist.gene.html
  rownames(percentDistance) <- unique(snpTable$Genomes)
  colnames(percentDistance) <- unique(snpTable$Genomes)
  
  longdf <- melt(percentDistance, value.name = "GenD")
  longdf$Var1 <- factor(longdf$Var1, levels = levels)
  longdf$Var2 <- factor(longdf$Var2, levels = levels)
  return(longdf)
}

GeoDistanceCalculator <- function(metadata, levels) {
  metadataFiltered <- metadata %>%
  filter(Name %in% levels)
  
  lonlatmatrix <- metadataFiltered %>%
    select(Longitud, Latitude) %>%
    data.matrix()
  distanceMatrix <- distm(lonlatmatrix, fun = distCosine)
  rownames(distanceMatrix) <- metadataFiltered$Name
  colnames(distanceMatrix) <- metadataFiltered$Name
  longdf <- melt(distanceMatrix, value.name = "GeoD")
  longdf$Var1 <- factor(longdf$Var1, levels = levels)
  longdf$Var2 <- factor(longdf$Var2, levels = levels)
  return(longdf)
}

C14DistanceCalculator <- function(metadata, levels) {
  metadataFiltered <- metadata %>%
  filter(Name %in% levels)
  Median.C14 <- metadataFiltered$Median.C14
  medianDateDiff <- outer(Median.C14, Median.C14, '-')
  #medianDateDiff <- abs(medianDateDiff)
  rownames(medianDateDiff) <- metadataFiltered$Name
  colnames(medianDateDiff) <- metadataFiltered$Name
  longdf <- melt(medianDateDiff, value.name = "C14_distance")
  longdf$Var1 <- factor(longdf$Var1, levels = levels)
  longdf$Var2 <- factor(longdf$Var2, levels = levels)
  
  return(longdf)
}

matrixPlotting <- function(df, fi, title) {
  ggplot(df, aes(x= Var2, y=Var1))+
  geom_raster(aes(fill=fi)) +
  scale_fill_viridis_c(option="plasma") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Samples", y="Samples", title=title)
}

plotGenTime <- function(DNADist, GeoDist, TimeDist) {
  GenGeoTimeD <- DNADist %>%
  left_join(GeoDist) %>%
  left_join(TimeDist) %>%
  filter(C14_distance > 0)

lmGenTime <- lm(GenD ~ C14_distance, data = GenGeoTimeD)
GenTimeCorr <- ggplot(GenGeoTimeD, aes(x= GenD, y=C14_distance, color=GeoD)) +
  geom_point() +
  geom_smooth(method = lm) +
  scale_color_viridis_c(option = "plasma") +
  labs(title = paste("Adj R2 =", round(summary(lmGenTime)$adj.r.squared, digits = 4), ",p-value =", summary(lmGenTime)$coef[2,4])) +
  theme_bw()
return(GenTimeCorr)
}

plotGenGeo <- function(DNADist, GeoDist, TimeDist) {
  GenGeoTimeD <- DNADist %>%
  left_join(GeoDist) %>%
  left_join(TimeDist) %>%
  filter(C14_distance > 0)

lmGenGeo <- lm(GenD ~ GeoD, data = GenGeoTimeD)
GenGeoCorr <- ggplot(GenGeoTimeD, aes(x= GenD, y=GeoD, color=C14_distance)) +
  geom_point() +
  geom_smooth(method = lm) +
  scale_color_viridis_c(option = "plasma") +
  labs(title = paste("Adj R2 =", round(summary(lmGenGeo)$adj.r.squared, digits = 4), ",p-value =", summary(lmGenGeo)$coef[2,4])) +
  theme_bw()
return(GenGeoCorr)
}
###END FUNCTIONS ###
```
###LNBA branch
```{r}
##Reading metadata file
metadata <- read.csv("Data/2020-07-09_LNBA_leprosy_enterica_comp/LNBA_transect/Metadata_coordinates_dating_sex_updated_def.csv") %>%
  filter(!Name=="XXX001.A18956.61")

##Reading snpTable and selecting genomes present in the metadata file
snpTable <- read.delim("Data/2020-07-09_LNBA_leprosy_enterica_comp/LNBA_transect/genotyped_snpTable_OOH003_c.tsv") %>%
  pivot_longer(c(-Position, -Ref), names_to = "Genomes", values_to = "Call") %>%
  filter(Genomes %in% metadata$Name) %>%
  mutate(CallDef = ifelse(Call == ".", as.character(Ref), as.character(Call))) %>%
  select(-Ref, -Call) %>%
    filter(!Genomes=="XXX001.A18956.61")

missingData <- snpTable %>%
  filter(CallDef == "N") %>%
  group_by(Position) %>%
  summarise(percent=(n()/length(unique(snpTable$Genomes)))*100)

GenomesPresent <- unique(snpTable$Genomes) 
#Prepare data for calculating the genetic distance and excluding positions where all the genomes have the same call
snpDataPestis <- snpTable %>%
  pivot_wider(names_from = Position, values_from = CallDef) %>%
  select_if(~ length(unique(.)) > 1) %>%
  as.matrix() %>%
  as.DNAbin()

levelsPhylogeny <- metadata$Name

longDistDNAPercent <- GenDistanceCalculator(snpDataPestis, "raw", TRUE, levelsPhylogeny, snpTable)

#Calculate distance matrix
longDistMa <- GeoDistanceCalculator(metadata, levelsPhylogeny)

#Difference median C14
medianDateDiffLong <- C14DistanceCalculator(metadata, levelsPhylogeny)


GenTimeCorr <- plotGenTime(longDistDNAPercent, longDistMa, medianDateDiffLong)
GenGeoCorr <- plotGenGeo(longDistDNAPercent, longDistMa, medianDateDiffLong)

plotLNBA <- ggarrange(GenTimeCorr, GenGeoCorr, ncol = 2)
plotLNBAannotated <- annotate_figure(plotLNBA, top= text_grob("Y. pestis LNBA branch", face = "italic"))

##Convert to function
#plot_ly(data = GenGeoTimeD, x= ~GenD, y= ~C14_distance, z= ~GeoD, type = "scatter3d", mode="markers", color = ~Var2)
```
###Second Pandemic Branch
```{r}
##Reading metadata file
metadata <- read.csv("Data/2020-07-09_LNBA_leprosy_enterica_comp/2ndPandemic_transect/Metadata_2ndpandemic.csv")

##Reading snpTable and selecting genomes present in the metadata file
snpTable <- read.delim("Data/2020-07-09_LNBA_leprosy_enterica_comp/LNBA_transect/genotyped_snpTable_OOH003_c.tsv") %>%
  pivot_longer(c(-Position, -Ref), names_to = "Genomes", values_to = "Call") %>%
  filter(Genomes %in% metadata$Name) %>%
  mutate(CallDef = ifelse(Call == ".", as.character(Ref), as.character(Call))) %>%
  select(-Ref, -Call) %>%
    filter(!Genomes=="XXX001.A18956.61_merged")

missingData <- snpTable %>%
  filter(CallDef == "N") %>%
  group_by(Position) %>%
  summarise(percent=(n()/length(unique(snpTable$Genomes)))*100)

GenomesPresent <- unique(snpTable$Genomes) 
#Prepare data for calculating the genetic distance and excluding positions where all the genomes have the same call
snpDataPestis <- snpTable %>%
  pivot_wider(names_from = Position, values_from = CallDef) %>%
  select_if(~ length(unique(.)) > 1) %>%
  as.matrix() %>%
  as.DNAbin()

levelsPhylogeny <- metadata$Name

longDistDNAPercent <- GenDistanceCalculator(snpDataPestis, "raw", TRUE, levelsPhylogeny, snpTable)

#Calculate distance matrix
longDistMa <- GeoDistanceCalculator(metadata, levelsPhylogeny)

#Difference median C14
medianDateDiffLong <- C14DistanceCalculator(metadata, levelsPhylogeny)

GenTimeCorr <- plotGenTime(longDistDNAPercent, longDistMa, medianDateDiffLong)
GenGeoCorr <- plotGenGeo(longDistDNAPercent, longDistMa, medianDateDiffLong)

plot2ndpandemic <- ggarrange(GenTimeCorr, GenGeoCorr, ncol = 2)
plot2ndpandemicannotated <- annotate_figure(plot2ndpandemic, top= text_grob("Y. pestis second pandemic branch", face = "italic"))

##Convert to function
#plot_ly(data = GenGeoTimeD, x= ~GenD, y= ~C14_distance, z= ~GeoD, type = "scatter3d", mode="markers", color = ~Var2)

```
###S. enterica
```{r}
##Reading metadata file
metadata <- read.csv("Data/2020-07-09_LNBA_leprosy_enterica_comp/enterica/Enterica_metadata.csv")

##Reading snpTable and selecting genomes present in the metadata file
snpTable <- read.delim("Data/2020-07-09_LNBA_leprosy_enterica_comp/enterica/snpTable.tsv") %>%
  pivot_longer(c(-Position, -Ref), names_to = "Genomes", values_to = "Call") %>%
  filter(Genomes %in% metadata$Name) %>%
  mutate(CallDef = ifelse(Call == ".", as.character(Ref), as.character(Call))) %>%
  select(-Ref, -Call)

#missingData <- snpTable %>%
#  filter(CallDef == "N") %>%
#  group_by(Position) %>%
#  summarise(percent=(n()/length(unique(snpTable$Genomes)))*100)

GenomesPresent <- unique(snpTable$Genomes) 
#Prepare data for calculating the genetic distance and excluding positions where all the genomes have the same call
snpData <- snpTable %>%
  pivot_wider(names_from = Position, values_from = CallDef) %>%
  select_if(~ length(unique(.)) > 1) %>%
  as.matrix() %>%
  as.DNAbin()

#To order the matrix by phylogeny
levelsPhylogeny <- metadata$Name

longDistDNAPercent <- GenDistanceCalculator(snpData, "raw", TRUE, levelsPhylogeny, snpTable)

#Calculate distance matrix
longDistMa <- GeoDistanceCalculator(metadata, levelsPhylogeny)

#Difference median C14
medianDateDiffLong <- C14DistanceCalculator(metadata, levelsPhylogeny)

GenTimeCorr <- plotGenTime(longDistDNAPercent, longDistMa, medianDateDiffLong)
GenGeoCorr <- plotGenGeo(longDistDNAPercent, longDistMa, medianDateDiffLong)

plotEnterica <- ggarrange(GenTimeCorr, GenGeoCorr, ncol = 2)
plotEntericaAnnotated <- annotate_figure(plotEnterica, top= text_grob("S. enterica", face = "italic"))
```
###M. leprae
```{r}
##Reading metadata file
metadata <- read.csv("Data/2020-07-09_LNBA_leprosy_enterica_comp/leprosy/Leprosy_metadata.csv")

##Reading snpTable and selecting genomes present in the metadata file
snpTable <- read.delim("Data/2020-07-09_LNBA_leprosy_enterica_comp/leprosy/2018_02_12_snpTable_wo_hyper.tsv") %>%
  pivot_longer(c(-Position, -Ref), names_to = "Genomes", values_to = "Call") %>%
  filter(Genomes %in% metadata$Name) %>%
  mutate(CallDef = ifelse(Call == ".", as.character(Ref), as.character(Call))) %>%
  select(-Ref, -Call)

#missingData <- snpTable %>%
#  filter(CallDef == "N") %>%
#  group_by(Position) %>%
#  summarise(percent=(n()/length(unique(snpTable$Genomes)))*100)

GenomesPresent <- unique(snpTable$Genomes) 
#Prepare data for calculating the genetic distance and excluding positions where all the genomes have the same call
snpData <- snpTable %>%
  pivot_wider(names_from = Position, values_from = CallDef) %>%
  select_if(~ length(unique(.)) > 1) %>%
  as.matrix() %>%
  as.DNAbin()

#To order the matrix by phylogeny
levelsPhylogeny <- metadata$Name

longDistDNAPercent <- GenDistanceCalculator(snpData, "raw", TRUE, levelsPhylogeny, snpTable)

#Calculate distance matrix
longDistMa <- GeoDistanceCalculator(metadata, levelsPhylogeny)

#Difference median C14
medianDateDiffLong <- C14DistanceCalculator(metadata, levelsPhylogeny)

GenTimeCorr <- plotGenTime(longDistDNAPercent, longDistMa, medianDateDiffLong)
GenGeoCorr <- plotGenGeo(longDistDNAPercent, longDistMa, medianDateDiffLong)

plotLeprosy <- ggarrange(GenTimeCorr, GenGeoCorr, ncol = 2)
plotLeprosyannotated <- annotate_figure(plotLeprosy, top= text_grob("Mycobacterium Leprae", face = "italic"))
```

```{r}
ggarrange(plotLNBAannotated, plot2ndpandemicannotated, plotEntericaAnnotated, plotLeprosyannotated,
          nrow = 4,
          labels = c("A","B","C", "D"))
```


#Indels
```{r}
library(tidyverse)
library(gridExtra)
library(ggplot2)
library(ggalt)

plottingMissingRegions <- function(df, levelvector, genomeToFilter, lengthChromosome, titlePlot){
  msdf <- df
  colnames(msdf) <-  c('Genome','PositionStart','PositionEnd','CoverageOther', 'Length','PercentCov')
  msdf$Genome2 <- factor(msdf$Genome, levels = levelvector)
  msdf %>% filter(! Genome %in% genomeToFilter) %>%
  ggplot(aes(x=PositionStart, xend=PositionEnd, y=Genome2, group=Genome)) +
    geom_dumbbell(size=1.5, color="#a0117d", colour_x = "#ffd23f", colour_xend = "#a29f15") +
    scale_x_continuous(expand = c(0,0), limits = c(0,lengthChromosome)) +
    ggtitle(titlePlot) +
    labs(x = "Postion Chromosome") +
    theme_minimal() +
    theme(axis.title.y = element_blank())
}

plottingMissingRegions2 <- function(df, levelvector, genomeToFilter, lengthChromosome, titlePlot){
  msdf <- df
  colnames(msdf) <-  c('Genome','PositionStart','PositionEnd', 'Length','PercentCov')
  msdf$Genome2 <- factor(msdf$Genome, levels = levelvector)
  msdf %>% filter(! Genome %in% genomeToFilter) %>%
  ggplot(aes(x=PositionStart, xend=PositionEnd, y=Genome2, group=Genome)) +
    geom_dumbbell(size=2, color="#2804c9", colour_x = "#2804c9", colour_xend = "#2804c9") +
    scale_x_continuous(expand = c(0,0),limits = c(0,lengthChromosome)) +
    ggtitle(titlePlot) +
    labs(x = "Postion Chromosome") +
    theme_minimal() +
    theme(axis.title.y = element_blank())
}

missingregionsCO92_merged <- read.delim("/projects1/pestis/lnba_paper_2020/indels/CO92_based_eager2_indelScript/Missingregions_min500bp_100.tsv", header = F) %>%
  mutate(V1 = str_replace(V1, ".reads.fastq.sorted.cleaned.rmdup.MarkD_", "")) %>%
  mutate(V1 = str_replace(V1, ".reads.fastq.sorted.cleaned.rmdup.MarkD", "")) %>%
  mutate(V1 = str_replace(V1, ".merged.samereadgroup.cleaned", "")) %>%
  mutate(V1 = str_replace(V1, ".trim.fastq.mapq0.onlymapped_","")) %>%
  mutate(V1 = str_replace(V1, "_L1","")) %>%
  mutate(V1 = str_replace(V1, "0.PE2-PESTF_0.1", "0.PE2_PEST-F"))

missingregionsYpseudo_merged <- read.delim("/projects1/pestis/lnba_paper_2020/indels/ypseudo_based_eager2_indelScript/Missingregions_min500bp_100.tsv", header = F) %>%
  mutate(V1 = str_replace(V1, ".reads.fastq.sorted.cleaned.rmdup.MarkD_", "")) %>%
  mutate(V1 = str_replace(V1, ".reads.fastq.sorted.cleaned.rmdup.MarkD", "")) %>%
  mutate(V1 = str_replace(V1, ".samereadgroup.cleaned", "")) %>%
  mutate(V1 = str_replace(V1, ".trim.fastq.sorted.cleaned.rmdup.MarkD_","")) %>%
  mutate(V1 = str_replace(V1, "_L1",""))


Order <- c("Y.pseudo.0.1_32.fq", "RISE509.MA877", "RK1001.C0101_allcombined","VLI092.A0101_rmdup", "KNK001_libmerged", "GEN_72", "GRS004.A0101_rmdup", "ABI745_HiSeqL5L6", "XXX001.A18962-69-72_libmerged", "GLZ001.A0101_02.YP2.1", "GLZ002.A0101_2.YP2.1", "ABI843_nonUDG", "ABI933_1343UNTA85", "HOP001.A0102_rmdup", "HOP004_libmerged", "KLZ001_libmerged","CHC004.A0101_rmdup","ABI910_6POST", "HGC009_libmerged", "KLE031.A0101_rmdup", "MIB054_libmerged", "KLE048.A0101_rmdup", "OOH003.A0102_A0103.merged", "RISE505.MA873",  "ARS007_libmerged", "KZL002_libmerged", "0.PE2_PEST-F", "XXX001.A18956-61_libmerged", "RT5_shotgun", "YpestisCO92")
toExclude <- c("HGC009_libmerged")
ggplot(missingregionsCO92_merged, aes(V5)) + geom_histogram() + stat_bin(binwidth = 0.001) + geom_vline(aes(xintercept=0.15))

plottingMissingRegions2(missingregionsCO92_merged, Order, toExclude, 4653728, "Missing regions (reference = Y. pestis CO92)")
missingregionsCO92_merged %>%
  filter(V5 <= 0.20) %>%
  plottingMissingRegions2(Order, toExclude, 4744671, "Missing regions (reference = Y. pestis CO92)")
missingregionsCO92_merged %>%
  filter(V5 <= 0.15) %>%
  plottingMissingRegions2(Order, toExclude, 4744671, "Missing regions (reference = Y. pestis CO92)")


plottingMissingRegions2(missingregionsYpseudo_merged, Order, toExclude, 4744671, "Missing regions (reference = Y. pseudotuberculosis IP32953)")
missingregionsYpseudo_merged %>%
  filter(V5 <= 0.10) %>%
  plottingMissingRegions2(Order, toExclude, 4744671, "Missing regions (reference = Y. pseudotuberculosis IP32953)")


CO92_missing <- missingregionsYpseudo_merged %>%
  filter(V1=="YpestisCO92") %>%
  select(V2,V3,V4) %>%
  mutate(start_min=V2-60,
         start_max=V2+60,
         end_min=V3-60,
         end_max=V3+60,
         window_min=V4-60,
         window_max=V4+60)
ypseudo_trial <- missingregionsYpseudo_merged 
for(r in 1:nrow(CO92_missing)) {
  start_min<-CO92_missing[r,4]
  start_max<-CO92_missing[r,5]
  end_min<-CO92_missing[r,6]
  end_max<-CO92_missing[r,7]
  ypseudo_trial <- ypseudo_trial %>%
    filter(! V2 %in% (start_min:start_max))
}

plottingMissingRegions2(ypseudo_trial, Order, toExclude, 4744671, "Missing regions (reference = Y. pseudotuberculosis IP32953)")


```

#Missing genes
```{r}
missinggenes <- read.delim("/projects1/pestis/lnba_paper_2020/indels/CO92_based/All.missing.genes.bed")
colnames(missinggenes) <- c("Genome", "PositionStart", "PositionEnd", "Randomtab", "Genes", "BasesMissing")

missinggenes <- missinggenes %>% select(-Randomtab) #%>%
#  mutate(Clade=)
#genes <- unique(missinggenes$Genes)
#write.csv(genes, "/projects1/pestis/lnba_paper_2020/indels/CO92_based/unique_genes.txt", row.names = F, col.names = F)
#Runned in the cluster in /projects1/pestis/lnba_paper_2020/indels/CO92_based
#system("tail -n +2 unique_genes.txt | sed 's/"//g' > unique_genes_def.txt")
#system("rm unique_genes.txt")
#system("grep -f unique_genes_def.txt Y_pestis_NC_003143_genes.bed | awk -F '\t' '{print $NF"\t"($3-$2)}' > genes_cdssize.tsv")

genescds <- read.delim("/projects1/pestis/lnba_paper_2020/indels/CO92_based/genes_cdssize.tsv", header = F)
colnames(genescds) <- c("Genes","CDS_size")

missinggenesdef <-missinggenes %>%
  left_join(genescds) %>%
  group_by(Genes, Genome, CDS_size) %>%
  summarise(BasesMissing=sum(BasesMissing)) %>%
  mutate(PercentGenePresent=((CDS_size-BasesMissing))/CDS_size)

missinggenesdef %>%
  filter(PercentGenePresent<0.90) %>%
  ggplot(aes(PercentGenePresent)) +                     # Plot the values
    geom_histogram(bins = 300, aes(fill=..count..)) +
  scale_fill_gradient("Count", low="green", high="red")
LNBAGenomes <- c( "RISE509.MA877.trim", "RK1001.C0101_allcombined", "KNK001.A0101.YP1", "GEN_72", "GRS004.A0101.YP1", "ABI745_HiSeqL5L6", "XXX001.A18962-69.merged", "XXX001.A18962-73.merged", "XXX001.A18970-73.merged","XXX001.A18962-69-72.merged", "ABI843_nonUDG", "ABI933_1343UNTA85", "HOP001.A0102_trimmed.combined", "HOP004_A0101", "KLZ001.A0101.YP1",  "ABI910_6POST", "HGC009.A0101.YP1", "KLE031.A0101_fqconcatenated", "KLE048.A0101_fqconcatenated", "RISE505.MA873_L1", "OOH003_A0102",  "ARS007_B0101", "KZL002_A0101")


summary_missing_LNBA <- missinggenesdef %>%
  mutate(Clade=ifelse(Genome %in% LNBAGenomes, "LNBA", "NoLNBA")) %>%
  filter(PercentGenePresent<0.75) %>%
  filter(! Genome %in% c("KNK001.A0101.YP1","HGC009.A0101.YP1","XXX001.A18962-69.merged","XXX001.A18970-73.merged","XXX001.A18962-73.merged", "MIB054.A0101.YP1")) %>%
  group_by(Genes,Clade) %>%
  summarise(total=n())
summary_missing_LNBA %>%
  ggplot(aes(total)) +
  geom_histogram(bins = 20, aes(fill=..count..)) +
  scale_fill_gradient("Count", low="green", high="red")


noLNBA <- summary_missing_LNBA %>%
  filter(Clade=="NoLNBA")
LNBA <- summary_missing_LNBA %>%
  filter(Clade=="LNBA")
Genes3ormore <- summary_missing_LNBA %>%
  filter(Clade=="LNBA") %>%
  filter(total>=3) %>%
  filter(! Genes %in% noLNBA$Genes)
Genesin2 <- summary_missing_LNBA %>%
  filter(Clade=="LNBA") %>%
  filter(total==2) %>%
  filter(! Genes %in% noLNBA$Genes)
Singletons <- summary_missing_LNBA %>%
  filter(Clade=="LNBA") %>%
  filter(total==1) %>%
  filter(! Genes %in% noLNBA$Genes)

NoinLNBA <- noLNBA %>%
  filter(! Genes %in% LNBA$Genes)

missinggenesdef$Genome2 <- factor(missinggenesdef$Genome, levels = c("Y.pseudo.0", "RISE509.MA877.trim", "RK1001.C0101_allcombined", "KNK001.A0101.YP1", "GEN_72", "GRS004.A0101.YP1", "ABI745_HiSeqL5L6", "XXX001.A18962-69.merged", "XXX001.A18962-73.merged", "XXX001.A18970-73.merged","XXX001.A18962-69-72.merged", "ABI843_nonUDG", "ABI933_1343UNTA85", "HOP001.A0102_trimmed.combined", "HOP004_A0101", "KLZ001.A0101.YP1",  "ABI910_6POST", "HGC009.A0101.YP1", "KLE031.A0101_fqconcatenated", "KLE048.A0101_fqconcatenated", "RISE505.MA873_L1", "OOH003_A0102",  "ARS007_B0101", "KZL002_A0101" ,"XXX001.A18956-61.merged","0.PE2-PESTF_0.1"))
#missinggenesdef$Genes2 <- factor(missinggenesdef$Genes, levels = as.factor(genescds$Genes))
plottingHist <- function(df, df2) {
  df %>%
    filter(Genes %in% df2$Genes) %>%
  filter(! Genome %in% c("KNK001.A0101.YP1","HGC009.A0101.YP1", "MIB054.A0101.YP1")) %>%
  mutate(PercentGenePresent2=ifelse(PercentGenePresent<0, 0, PercentGenePresent)) %>%
  ggplot(aes(Genes, Genome2, fill=PercentGenePresent2)) +
  geom_tile() +
  scale_fill_viridis_c() + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 90, size = 5), legend.title = element_blank(), panel.background = element_rect(fill = '#FDE725FF'))
}
plottingHist(missinggenesdef, Genes3ormore)
plottingHist(missinggenesdef, Genesin2)
plottingHist(missinggenesdef, Singletons)
plottingHist(missinggenesdef, NoinLNBA)
```

#SNPEff
```{r}
library(tidyverse)
snpEffTable <- read.delim("/projects1/pestis/lnba_paper_2020/snpEff/snpEff_AAV_3X_snp_falseSNPsExcluded_historicalgenomes10X_OOH003snpcalling_XXX001.A18962-69-72_2020-07-20/snpEff_lnba_20200721_2.out", skip = 2)
snpTableOOH003c <- read.delim("/projects1/pestis/lnba_paper_2020/vcfAnalysis/AAV_3X_snp_falseSNPsExcluded_historicalgenomes10X_OOH003snpcalling_XXX001.A18962-69-72_2020-07-20/genotyped_snpTable_OOH003_c.tsv", check.names=FALSE)

snpTableOOH003c %>%
  left_join(snpEffTable) %>%
  select(-X..Chromo, -Change, -Reference) %>%
  write_tsv("/projects1/pestis/lnba_paper_2020/snpEff/snpEff_AAV_3X_snp_falseSNPsExcluded_historicalgenomes10X_OOH003snpcalling_XXX001.A18962-69-72_2020-07-20/snpTable_OOH003genotyped_withSNPEff.tsv")

```
#Figure displaying the radiocarbon range per sample
```{r}
metadataNBA <- read.csv("Data/Radiocarbon_dates/Metadata_coordinates_dating_sex.csv")

metadataNBA$Sample.Name <- factor(metadataNBA$Sample.Name, levels= c("RISE509", "RK1001", "VLI092.A0101","KNK001","GEN72","GRS004.A0101","Gyvakaral1","I5884","GLZ002","GLZ001","KunilaII","1343UnTal85","HOP004", "HOP001.A0102","KLZ001","CHC004.A0101","6Post","KLE031.A0101","MIB054", "KLE048.A0101","OOH003","RISE505","ARS007", "KZL002","I2470","MIK005_RT5"))

library(ggalt)
ggplot() +
  geom_dumbbell(data = metadataNBA, aes(y=Sample.Name, x=Start_Range_C14, 
                                        xend=End_Range_C14, colour=Colour), size=1, show.legend = FALSE) +
  scale_x_reverse() +
  xlab("cal. years Before Present") +
  ylab("") +
  theme_bw()
```
